# GM - ColorGroup/Raw Material

Elliptical bifaces from the Millsap Cache and Jowell Farm were used to examine whether their morphology remains stable, or expresses morphological variability by ColorGroup/raw meterial. Variability in elliptical biface morphology provides evidence associated with differences in morphology as a function of geography/raw material.

This analysis follows recent analyses of Perdiz arrow points ([Selden and Dockall, in review](https://github.com/seldenlab/perdiz3/blob/main/ms/perdiz3/perdiz3.pdf)) and Gahagan biface morphology across the southern Caddo area and central Texas [@RN11783;@RN21001;@RN11264], where Perdiz arrow points and Gahagan bifaces have been found to differ in shape across the same geography as Caddo bottles [@RN11801;@RN11782;@RN11716;@RN8312]. This analysis also follows another geometric morphometric analysis in which Perdiz arrow points from the ancestral Caddo area were found to differ as a function of _time_, _raw material_, and _burial context_ [@RN9551].

## Load packages + data

```{r load.packagesColorGroupraw, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# load required analysis packages
library(here)
library(StereoMorph)
library(geomorph)
library(tidyverse)
library(wesanderson)
library(ggplot2)

# read data and define number of sLMs ----
shapes <- readShapes("shapes")
shapesGM <- readland.shapes(shapes, nCurvePts = c(8, 10, 18))

# read qualitative data
qdata <- read.csv("qdata.csv", 
                  header = TRUE, 
                  row.names = 1)
```

## Generalised Procrustes Analysis

Landmark data were aligned to a global coordinate system [@RN11622;@RN11623;@RN11563], achieved through generalized Procrustes superimposition [@RN478] performed in R 4.1.3 [@R] using the `geomorph` library v. 4.0.3 [@RN11530;@RN1774;@RN9565]. Procrustes superimposition translates, scales, and rotates the coordinate data to allow for comparisons among objects [@RN11564;@RN478]. The `geomorph` package uses a partial Procrustes superimposition that projects the aligned specimens into tangent space subsequent to alignment in preparation for the use of multivariate methods that assume linear space [@RN1646;@RN11563]. 

```{r outlinesColorGroupraw, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# gpa ----
Y.gpa <- gpagen(shapesGM, print.progress = FALSE)

# plot
plot(Y.gpa)

# geomorph data frame ----
gdf <- geomorph.data.frame(shape = Y.gpa$coords, 
                           size = Y.gpa$Csize,
                           ColorGroup = qdata$ColorGroup)

# add centroid size to qdata
qdata$csz <- Y.gpa$Csize
```

## Principal Components Analysis

Principal components analysis [@RN1746] was used to visualise shape variation among the Perdiz arrow points. The shape changes described by each principal axis are commonly visualized using thin-plate spline warping of a reference 3D mesh [@RN1731;@RN479].


```{r pca.plotColorGroupraw, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# principal components analysis ----
pca<-gm.prcomp(Y.gpa$coords)
summary(pca)

# set plot parameters
pch.gps <- c(1,3)[as.factor(qdata$ColorGroup)]
col.gps <- wes_palette("Moonrise2")[as.factor(qdata$ColorGroup)]
col.hull <- c("#798E87","#C27D38")

# plot pca by merged
pc.plot <- plot(pca, 
                asp = 1,
                pch = pch.gps,
                col = col.gps)
shapeHulls(pc.plot, 
           groups = qdata$ColorGroup,
           group.cols = col.hull)
```

## Procrustes ANOVA

To assess whether shape and size differ by `size class`, Procrustes ANOVAs [@RN1749] were run that enlist effect-sizes (zscores) computed as standard deviates of the generated sampling distributions [@RN1756]. A residual randomization permutation procedure (RRPP; n = 10,000 permutations) was used for all Procrustes ANOVAs [@RN1655;@RN11775], which has higher statistical power and a greater ability to identify patterns in the data should they be present [@RN1719].


```{r anovaColorGroupraw, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# ANOVA ----
# shape
fit.sh <- procD.lm(shape ~ ColorGroup,
                   data = gdf,
                   print.progress = FALSE,
                   iter = 9999)

# shape as a function of ColorGroup/raw material?
anova(fit.sh)

# size
fit.sz <- procD.lm(size ~ ColorGroup,
                   data = gdf,
                   print.progress = FALSE,
                   iter = 9999)

# size as a function of ColorGroup/raw material?
anova(fit.sz)
```
