--- 
title: "A quantitative assessment of Jowell knife morphology"
author: 
    - Robert Z. Selden, Jr.^[Heritage Research Center, Stephen F. Austin State University; Department of Biology, Stephen F. Austin State University; and Cultural Heritage Department, Jean Monnet University, zselden@sfasu.edu]
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
description: "Pilot study of Jowell knife morphology."
---

# Preface {-}

Placeholder


## Primary findings
## Acknowledgments
## Funding {-}
## Data management
## Colophon

<!--chapter:end:index.Rmd-->


# Linear metrics by site

Placeholder


## Set working directory, load data, and define variables
## Boxplots for _variable_ by _site_
## Ternary plot
## Principal Components Analysis for _site_
## Analyses of Variance (ANOVA) for _variable_ ~ _site_
### _Maximum length_ of Jowell knives
### _Maximum width_ of Jowell knives
### _Maximum thickness_ of Jowell knives

<!--chapter:end:01-linear.Rmd-->

# Geometric morphometrics 

## Load packages + data

```{r load.packages, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# install required analysis packages
#devtools::install_github("r-lib/here")
#devtools::install_github("karthik/wesanderson")
#devtools::install_github("MomX/Momocs")

# load packages
library(here)
library(wesanderson)
library(ggplot2)
library(Momocs)

# read images + attribute data
jpg.list <- list.files(here("./jpegs"), full.names = TRUE)
att.data <- read.csv("att.data.csv", header = TRUE, as.is = TRUE)

# attributes to factors
att.data$site <- as.factor(att.data$site)
```

## Generate outlines

```{r outlines, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# generate outlines
outlines <- jpg.list %>%
  import_jpg()

# add attributes
data.out <- Out(outlines, 
         fac = att.data)

# center, scale, align, and rotate specimens
norm.outlines <- data.out %>% 
  coo_center %>%
  coo_scale() %>%
  coo_align() %>% 
  coo_rotate()

## 41an13
an13 <- filter(norm.outlines, 
                 site %in% "41AN13")

an13 <- an13 %>% 
  coo_scale() %>%
  coo_align() %>%
  coo_rotate() %>% 
  coo_center()

## 41hs261
hs261 <- filter(norm.outlines, 
                 site %in% "41HS261")

hs261 <- hs261 %>% 
  coo_scale() %>%
  coo_align() %>%
  coo_rotate() %>% 
  coo_center()

# render figure
par(mfrow=c(2, 1))

stack(an13, 
      title = "41AN13", 
      xy.axis = TRUE, 
      ldk = FALSE, 
      centroid = FALSE, 
      meanshape = TRUE, 
      meanshape_col = "blue")

stack(hs261, 
      title = "41HS261", 
      xy.axis = TRUE, 
      ldk = FALSE, 
      centroid = FALSE, 
      meanshape= TRUE, 
      meanshape_col = "blue")
```

## Calibrate harmonic + EFA

```{r cal.harm, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# calibrate how many harmonics needed
calibrate_harmonicpower_efourier(norm.outlines, nb.h = 30)

# 9 harmonics needed to capture 99 percent of variation
calibrate_reconstructions_efourier(norm.outlines, range = 1:9)

# generate efa outlines with 9 harmonics
efa.outlines <- efourier(norm.outlines, nb.h = 9, norm = TRUE)

# use efa.outlines for pca
pca.outlines <- PCA(efa.outlines)
```

## PCA

```{r pca.plot, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# pca 
scree_plot(pca.outlines)

# plot pca by site
plot_PCA(pca.outlines, 
         morphospace_position = "range_axes",
         ~site, 
         zoom = 1.25)

# contribution of each pc
boxplot(pca.outlines, 
        ~site, 
        nax = 1:5)

# mean shape + 2sd for the first 10 pcs
PCcontrib(pca.outlines, 
          nax = 1:5)
```

## MANOVA

```{r manova, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# manova
# does shape differ between sites?
MANOVA(pca.outlines, 'site')
```

## Mean shapes

```{r ms1, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE, fig.cap="Pairwise comparison of mean shapes for Jowell knives by site."}
# mean shapes
ms.1 <- MSHAPES(efa.outlines, ~site)
plot_MSHAPES(ms.1, size = 0.75)
```

<!--chapter:end:02-2dgm.Rmd-->

`r if (knitr::is_html_output()) '
# References {-}
'`

<!--chapter:end:06-references.Rmd-->

